Time Series Modeling of Varying Complexity Captures Subdiffusive Solute Dynamics and Predicts Long Timescale Behavior in Nanoscale Pores.

### 2019-12-13 Ben_Manuscripts/stochastic_transport/Draft.tex
Incorporated MSDDM into MFPT calculations

  The models are capable of reasonably predicting the MD MSD values of the second
  half of the solute trajectories based on parameters generated from the first half
  when the dwell time distributions are parameterized by a power law with an 
  exponential cut-off. Without truncation of the power law distribution, the MD 
  MSDs are underestimated in all cases because dwell times on the order of the
  MD simulation length are sampled and incorporated into the simulated anomalous
  diffusion trajectories. Considering the longest observed dwell time among all
  solutes was 1.3 $\mu$s by ethylene glycol, we believe truncation is well-justified.

@@

  From a qualitative perspective, the models only have moderate success at predicting
  the shape of the MSD curves. In general, the MSDs generated by our AD models appear
  close to linear at long time scales. This is expected behavior for time averaged 
  MSDs, especially when truncating the power law distribution of dwell times.
  \cite{molina-garcia_crossover_2018} The linearity of Urea's MD MSD curve is 
  consistent with that of the AD model, but the MD MSD of acetic acid appears to have considerable
  curvature. Acetic acid's MSD is under-predicted at short timescales but within error 
  of MD (using sFBMcut and sFLMcut models) after a 500 ns time lag. The curvature suggests that the 
  model under-estimates long timescale positional anti-correlation which points to an
  over-estimate of the Hurst parameter. This is not surprising because the Hurst 
  parameter is challenging to quantify, especially with a relatively small amount of 
  data (see Section~\ref{method:sfbm} for further discussion of this challenge).
  A more accurate measurement of $H$ would fix the shape of the MSD curve, but also
  lower the predicted MSD meaning we are either underestimating the width of the hop
  length distribution, favoring longer dwell times or both.
  % BJC4: play around with this to see how to change shape of acetic acid curve. Hurst makes it arc more
  % ctrwsim.py -alpha 0.08 -fix_time -hop fbm -H 0.34 -ntraj 100 -dwell powerlaw_cutoff -lamb 0.0033 -hs 0.27

  This brief analysis suggests that working with only half of the data we collected
  ($\sim$2$\mu$s post-equilibration) is not always sufficient for extracting reliable
  parameter estimates. In most cases, the magnitude of the MSD predictions after a 
  500 ns time lag are within or close to within error of the MD MSDs (for sFBMcut and
  sFLMcut). However, in the case of acetic acid, this may only happen because the 
  Hurst parameter is overestimated. It appears we are operating on the border of the
  minimum amount of data required to accurately parameterize the AD model. Longer
  simulations and more independent trajectories may be necessary. Therefore, in 
  the next section we will work with parameters fit to the full equilibrated portion
  of the solute trajectories, doubling the data.

@@

  We used the 1-mode sFBMcut (see Table~\ref{table:anomalous_models}) anomalous 
  diffusion model in order to demonstrate how one can use its realizations in
  order to calculate the flux (section~\ref{method:mfpt}) of solutes given model 
  parameters extracted from MD simulations. The sFLMcut model and MSDDM yield 
  similar predictions at a much greater computational cost due to the slow
  generation of fractional L\'evy noise.
  %MRS5: One question is if other method would lead to substantially
  %different results for fluxes or selectivity.  May need too think
  %about justification more if just one method is mentioned.
  %BJC5: would love to test MSDDM against this but I don't know how to make 
  % the algorithm faster. It already uses an FFT to correlate the data (to which I pass powers of 2)

>>>>

  The models are capable of reasonably predicting the MD MSD values of the second
  half of the solute trajectories based on parameters generated from the first half
  when the dwell time distributions are parameterized by a power law with an 
  exponential cut-off. At long timescales, the MSD of Urea is under-predicted 
  for both the 1 and 2 mode models with the same true of acetic acid on short 
  timescales. Without truncation of the power law distribution, the MD 
  MSDs are underestimated in all cases because dwell times on the order of the
  MD simulation length are sampled and incorporated into the simulated anomalous
  diffusion trajectories. Considering the longest observed dwell time among all
  solutes was 1.3 $\mu$s by ethylene glycol, we believe truncation is well-justified.
  Although in most cases, the predicted MSD curves overlap the 1-$\sigma$ confidence
  intervals of the MD MSD curves, they appear to lie systematically below the mean. 
  It is not clear whether this is a consistent feature of the AD model since we 
  only used 2 solutes for this analysis.

@@

  From a qualitative perspective, the models only have moderate success at predicting
  the shape of the MSD curves. Error in the MSD curvature can help explain some of
  the error in the predicted MSD magnitudes. The curves predicted for Urea with both 1 
  and 2 modes appear to have too much curvature, which causes it to under-predict the 
  MSD at long timescales, while those of acetic acid lack curvature, leading the
  AD model to under-predict the MSD at short timescales.

  The under-estimate of Urea's MSD at long timescales is due to long timescale
  positional anti-correlation which may not be present in the real system. The 
  persistent curvature of the predicted MSD curves is a direct consequence of 
  the Hurst parameter. Without anti-correlation, the process would be 
  a pure CTRW for which one would expect the time averaged MSD curves to become 
  linear.~\cite{meroz_toolbox_2015} It may be true that on the $\mu$s time scale, 
  positional correlation is lost which would manifest as a transition from a sub-linear
  to linear MSD curve. A solution for more accurately modeling this behavior in  
  the future may be to truncate the positional autocorrelation function.~\cite{molina-garcia_crossover_2018}

  The under-estimate of the curvature of acetic acid's predicted MSD suggests that,
  in this case, the AD model over-estimates the Hurst parameter. This is not surprising
  because the Hurst parameter is challenging to quantify, especially with a relatively 
  small amount of data (see Section~\ref{method:sfbm} for further discussion of this 
  challenge). A more accurate measurement of $H$ would fix the shape of the MSD curve,
  but also lower the predicted MSD meaning we are either underestimating the width of
  the hop length distribution, favoring longer dwell times or both.

%  Urea's curves appear to have
%  close to the right amount of curvature at short timescales

%  In general, the MSDs generated by our AD models appear
%  close to linear at long time scales. This is expected behavior for time averaged 
%  MSDs, especially when truncating the power law distribution of dwell times.
%  \cite{molina-garcia_crossover_2018} The linearity of Urea's MD MSD curve is 
%  consistent with that of the AD model, but the MD MSD of acetic acid appears to have considerable
%  curvature.

%  Acetic acid's MSD is under-predicted at short timescales but within error 
%  of MD (using sFBMcut and sFLMcut models) after a 500 ns time lag. The curvature suggests that the 
%  model under-estimates long timescale positional anti-correlation which points to an
%  over-estimate of the Hurst parameter. This is not surprising because the Hurst 
%  parameter is challenging to quantify, especially with a relatively small amount of 
%  data (see Section~\ref{method:sfbm} for further discussion of this challenge).
%  A more accurate measurement of $H$ would fix the shape of the MSD curve, but also
%  lower the predicted MSD meaning we are either underestimating the width of the hop
%  length distribution, favoring longer dwell times or both.
  % BJC4: play around with this to see how to change shape of acetic acid curve. Hurst makes it arc more
  % ctrwsim.py -alpha 0.08 -fix_time -hop fbm -H 0.34 -ntraj 100 -dwell powerlaw_cutoff -lamb 0.0033 -hs 0.27

  This brief analysis suggests two shortcomings of the AD model. First, positional
  anti-correlation may diminish after a sufficiently long time lag, dependent on the 
  solute being studied. Second, it is difficult to reliably parameterize the Hurst
  parameter which is important for accurately describing the curvature of the solute
  MSDs.

  This analysis also suggests that working with only half of the data we collected
  ($\sim$2$\mu$s post-equilibration) is not always sufficient for extracting reliable
  parameter estimates. In most cases, the magnitude of the MSD predictions after a 
  500 ns time lag are within or close to within error of the MD MSDs (for sFBMcut and
  sFLMcut), but still appear to systematically under-predict the mean. We may be
  operating on the border of the minimum amount of data required to accurately 
  parameterize the AD model. Longer simulations and more independent trajectories may
  be necessary. Therefore, in the next section we will work with parameters fit to the
  full equilibrated portion of the solute trajectories, doubling the data.

%  However, in the case of acetic acid, this may only happen because the 
%  Hurst parameter is overestimated. It appears we are operating on the border of the
%  minimum amount of data required to accurately parameterize the AD model. Longer
%  simulations and more independent trajectories may be necessary. Therefore, in 
%  the next section we will work with parameters fit to the full equilibrated portion
%  of the solute trajectories, doubling the data.

@@

  We used the 1-mode sFBMcut (see Table~\ref{table:anomalous_models}) anomalous 
  diffusion model in order to demonstrate how one can use its realizations in
  order to calculate the flux (section~\ref{method:mfpt}) of solutes given model 
  parameters extracted from MD simulations. The sFLMcut AD model and MSDDM yield 
  similar predictions at a much greater computational cost due to the slow
  generation of fractional L\'evy noise.
  %MRS5: One question is if other method would lead to substantially
  %different results for fluxes or selectivity.  May need too think
  %about justification more if just one method is mentioned.
  %BJC5: would love to test MSDDM against this but I don't know how to make 
  % the algorithm faster. It already uses an FFT to correlate the data (to which I pass powers of 2)
